---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.path = "README/README-",
  message = FALSE, warning = FALSE, fig.asp = 1, fig.align = 'center'
)
```

DirStats
========

<!-- badges: start -->
[![License](https://img.shields.io/badge/license-GPL%20v3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![Travis build status](https://travis-ci.org/egarpor/DirStats.svg?branch=master)](https://travis-ci.org/egarpor/DirStats)
[![CRAN status](https://www.r-pkg.org/badges/version/DirStats)](https://cran.r-project.org/package=DirStats)
<!-- badges: end -->

## Overview

Currently implementing nonparametric kernel density estimation, bandwidth selection, and other utilities for analyzing directional data. Further nonparametric tools expected to be included in subsequent releases.

## Installation

Get the released version from CRAN:

```{r, install-CRAN, eval = FALSE}
# Install the package
install.packages("DirStats")

# Load package
library(DirStats)
```

Alternatively, get the latest version from GitHub:

```{r, install-GitHub, eval = FALSE}
# Install the package
library(devtools)
install_github("egarpor/DirStats")

# Load package
library(DirStats)
```

```{r, load, echo = FALSE}
# Load package
library(DirStats)
```

## Usage

The following are examples of the usage of the Bai et al. (1988)'s kernel density estimator, the cross-validatory bandwidth selectors in Hall et al. (1987), and the plug-in bandwidth selectors in García-Portugués (2013).

### Compute bandwidths

```{r, bws}
# Sample from a von Mises--Fisher on S^2
q <- 2
n <- 300
set.seed(42)
samp <- rbind(rotasym::r_vMF(n = n / 3, mu = c(rep(0, q), 1), kappa = 5),
              rotasym::r_vMF(n = n / 3, mu = c(rep(0, q), -1), kappa = 5),
              rotasym::r_vMF(n = n / 3, mu = c(1, rep(0, q)), kappa = 5))

# LCV bandwidth
(h_LCV <- bw_dir_lcv(data = samp)$h_opt)

# LSCV bandwidth
(h_LSCV <- bw_dir_lscv(data = samp)$h_opt)

# ROT bandwidth
(h_ROT <- bw_dir_rot(data = samp))

# Mixture fit, for AMI and EMI bandwidth selectors
fit_mix <- bic_vmf_mix(data = samp)

# AMI bandwidth
(h_AMI <- bw_dir_ami(data = samp, fit_mix = fit_mix))

# EMI bandwidth
(h_EMI <- bw_dir_emi(data = samp, fit_mix = fit_mix)$h_opt)
```

### Compute kernel density estimator

```{r, kde, fig.asp = 2/3}
# Compute kde
l <- 200
th <- seq(0, 2 * pi, l = l)
phi <- seq(0, pi, l = l / 2)
ang <- expand.grid(th = th, phi = phi)
x <- to_sph(th = ang$th, ph = ang$phi)
kde <- kde_dir(x = x, data = samp, h = h_EMI)

# Visualization
contour(x = th, y = phi, z = matrix(kde, nrow = l, ncol = l / 2),
        col = viridisLite::viridis(15), 
        levels = round(seq(min(kde), max(kde), length.out = 15), 4), 
        lwd = 2, xlab = expression(theta), ylab = expression(phi))
points(to_rad(samp), pch = 16)
```

## References

Bai, Z. D., Rao, C. R., and Zhao, L. C. (1988). Kernel estimators of density function of directional data. *Journal of Multivariate Analysis*, 27(1):24--39. <https://doi.org/10.1016/0047-259X(88)90113-3>

García-Portugués, E. (2013). Exact risk improvement of bandwidth selectors for kernel density estimation with directional data. *Electronic Journal of Statistics*, 7:1655--1685. <https://doi.org/10.1214/13-ejs821>

Hall, P., Watson, G. S., and Cabrera, J. (1987). Kernel density estimation with spherical data. *Biometrika*, 74(4):751--762.
<https://doi.org/10.1093/biomet/74.4.751>
